### 2017-11-12 Draft2/AsymptoticNotation.tex
Respond to feedback
Work

Typically, big-O notation is used to analyze the time or space complexity of a function. In order to highlight the benefit of growth arrays, however, I will use the $\sim$ relation to analyze complexity. This is because for certain operations, growth arrays are only better than dynamic arrays by a constant factor. For example, dynamic arrays might use roughly $2n$ space for appending $n$ items, while growth arrays would use roughly $n$ space. Even though growth arrays are clearly better in this regard, the big-O space complexity for both data structures would be the same, $O(n)$.

My goal is to be able to compare the coefficients of the highest-order terms in both expressions. For example, I would like to take the ratio $2n / n$, see that it is $2$, and conclude that dynamic arrays allocate roughly twice as much as growth arrays for large $n$. However, big-O notation does not support this.

@@

I will also introduce the cousins of $\sim$, $\FluteLess$ and $\FluteLeq$. $f \FluteLeq g$ if and only if $f \FluteLess g$ or $f \sim g$. $\FluteLess$ is defined as follows:
\begin{align*}
f \FluteLess g \Biconditional \ExprNToInfty \frac f g < 1
\end{align*}
\subsection{Properties}
\label{subsec:AsymptoticProperties}

Here, I define various properties of the $\sim$ relation that will be used in my proofs. These properties themselves are proved in the appendix.

@@

The following theorem states that $\sim$ is a transitive relation. This property is used implicitly by proofs that chain multiple $\sim$s in the form $f \sim g \sim h$, then conclude that $f \sim h$.

\begin{theorem}
\label{thm:Transitivity}
	If $f \sim g$ and $g \sim h$, then $f \sim h$.
\end{theorem}

@@

>>>>

Typically, big-O notation is used to analyze the time or space complexity of a function. In order to highlight the benefit of growth arrays, however, I will use the $\sim$ relation to analyze complexity. This is because growth arrays are only better than dynamic ones by a constant factor for certain operations. For example, dynamic arrays might use roughly $2n$ space for appending $n$ items, while growth arrays would use roughly $n$ space. Even though growth arrays are clearly better in this regard, the big-O space complexity for both data structures would be the same, $O(n)$.

My goal is to be able to compare the coefficients of the highest-order terms in both expressions. For example, I would like to take the ratio $2n : n$, see that it is $2$, and conclude that dynamic arrays allocate roughly twice as much as growth arrays for large $n$. However, big-O notation does not support this.

@@

I will also introduce two cousins of $\sim$, $\FluteLess$ and $\FluteLeq$. $f \FluteLeq g$ if and only if $f \FluteLess g$ or $f \sim g$. $\FluteLess$ is defined as follows:
\begin{align*}
f \FluteLess g \Biconditional \ExprNToInfty \frac f g < 1
\end{align*}
When reading these symbols, one says ``$f$ is asymptotically less than $g$'' or ``$f$ is asymptotically less than or equal to $g$''.

\subsection{Properties}
\label{subsec:AsymptoticProperties}

Here, I define various properties of the $\sim$ relation that later proofs will use. The properties themselves are proved in the appendix.

@@

The following theorem states that $\sim$ is a transitive relation. This property is used implicitly by proofs that chain multiple $\sim$ signs in the form $f \sim g \sim h$, and then conclude that $f \sim h$.

\begin{theorem}
\label{thm:SimIsTransitive}
	If $f \sim g$ and $g \sim h$, then $f \sim h$.
\end{theorem}

@@

The following theorem states that if two functions are asymptotic, then they can be used interchangeably in an asymptotic inequality.

\begin{theorem}
\label{thm:InterchangeableInInequality}
	If $f \sim \Tweak{f}$ and $\Tweak{f} \FluteLess \TweakSecond{f}$, then $f \FluteLess \TweakSecond{f}$. If $g \sim \TweakSecond{g}$ and $\Tweak{g} \FluteLess \TweakSecond{g}$, then $\Tweak{g} \FluteLess g$.
\end{theorem}

### 2017-11-12 Draft2/FieldsProperties.tex
Work

In subsequent sections, I will implement algorithms for both dynamic and growth arrays. In this section, I define fields and properties for these data structures which the algorithms will use. \textbf{Fields} are variables associated with an object that may be read from or written to. \textbf{Properties} are trivial, constant-time methods that do not change state.

If $\VarList$ is a dynamic array, then it is assumed to have the following fields:

@@

As a dynamic array, $\VarList$ is also given the following properties. {\HdrNote} For pseudocode, I will use $:=$ to denote a definition, and $=$ to denote an equivalence check. Functions that return boolean values are suffixed with ?.

@@

When a dynamic array is instantiated, the following code should run:

@@

If $\VarList$ is a growth array, then it is assumed to have the following fields:

\begin{itemize}
	\item $\VarList.\FieldHead$ - The \textbf{head} of $\VarList$. It returns the buffer we are currently adding items to.
	\item $\VarList.\FieldTail$ - The \textbf{tail} of $\VarList$. \textbf{Important note: The tail is a dynamic array.} It returns a \textit{dynamic array} of references to buffers that are already filled with items. The tail can be thought of as a two-dimensional array.\\
	{\HdrNote} It may seem strange for a growth array to use the very data structure it is replacing. As shown in Lemma \ref{lem:VarUsefulIsOLogN}, however, only $O(\log n)$ many references are appended to the tail. Thus, the extra copying and allocations the tail performs is minuscule compared to other work done by the growth array.
	\item $\VarList.\FieldSize$ - The number of items in $\VarList$.
	\item $\VarList.\FieldCapacity$ - The \textbf{capacity} of $\VarList$. It returns the maximum number of items $\VarList$ can hold before it must grow.
\end{itemize}

@@

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Returns whether $\VarList$ is empty}
		\State $\VarList.\FieldEmpty := \VarList.\FieldSize = 0$
		\State
		\LineComment{Returns whether $\VarList$ is full}
		\State $\VarList.\FieldFull := \VarList.\FieldSize = \VarList.\FieldCapacity$
		\State
		\LineComment{Returns the capacity of $\FieldHead$}
		\State $\VarList.\FieldHeadCapacity := \VarList.\FieldHead.\FieldLength$
		\State
		\LineComment{Returns the size of $\FieldHead$}
		\LineComment{\textbf{Rationale:} $\FieldCapacity - \FieldHeadCapacity$ is the total capacity of the buffers in $\FieldTail$.}
		\LineComment{Then, $\FieldSize - \left( \FieldCapacity - \FieldHeadCapacity \right)$ is the number of items that were added}
		\LineComment{after depleting the buffers in $\FieldTail$.}
		\State $\VarList.\FieldHeadSize := \VarList.\FieldSize - \left(\VarList.\FieldCapacity - \VarList.\FieldHeadCapacity\right)$
	\end{algorithmic}
\end{algorithm}

When a growth array is instantiated, the following code should run:

>>>>

In later sections, I will implement algorithms for both dynamic and growth arrays. In this section, I will define fields and properties for those algorithms to use. \textbf{Fields} are variables associated with an object that may be read from or written to. \textbf{Properties} are trivial, constant time methods that do not change state.

Let $\VarList$ be a dynamic array. I will give $\VarList$ the following fields:

@@

As a dynamic array, $\VarList$ is also given the following properties. ({\HdrNote} For pseudocode, I will use $:=$ to denote a definition, and $=$ to denote an equivalence check. Functions that return boolean values are suffixed with ?.)

@@

The following code should run when $\VarList$ is instantiated:

@@

Now, let $\VarList$ be a growth array. I will give it the following fields:

\begin{itemize}
	\item $\VarList.\FieldHead$ - The \textbf{head} of $\VarList$. It returns the buffer new items are appended to.
	\item $\VarList.\FieldTail$ - The \textbf{tail} of $\VarList$. (\textbf{Important note: The tail is a dynamic array.}) It returns a dynamic array of pointers to buffers that are already filled with items. The tail can be thought of as a two-dimensional array.\\
	({\HdrRemark} It may seem strange for a growth array to use the very data structure it is replacing. Lemma \ref{lem:VarUsefulIsOLogN} shows, however, that only $O(\log n)$ many pointers are appended to the tail. Thus, the extra copying and allocations the tail performs is minuscule compared to other work done by the growth array.)
	\item $\VarList.\FieldSize$ - The number of items in $\VarList$.
	\item $\VarList.\FieldCapacity$ - The \textbf{capacity} of $\VarList$. It returns the maximum number of items $\VarList$ can hold before it must grow.
\end{itemize}

@@

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Returns whether $\VarList$ is full}
		\State $\VarList.\FieldFull := \VarList.\FieldSize = \VarList.\FieldCapacity$
		\State
		\LineComment{Returns the capacity of $\FieldHead$}
		\State $\VarList.\FieldHeadCapacity := \VarList.\FieldHead.\FieldLength$
		\State
		\LineComment{Returns the size of $\FieldHead$}
		\LineComment{\textbf{Explanation:} $\FieldCapacity - \FieldHeadCapacity$ is the number of items in $\FieldTail$.}
		\LineComment{$\FieldSize - \left( \FieldCapacity - \FieldHeadCapacity \right)$ is the number of items not in $\FieldTail$ (meaning, in $\FieldHead$).}
		\State $\VarList.\FieldHeadSize := \VarList.\FieldSize - \left(\VarList.\FieldCapacity - \VarList.\FieldHeadCapacity\right)$
	\end{algorithmic}
\end{algorithm}

The following code should run when $\VarList$ is instantiated:
